Data Management Plan

### 2021-11-30 application/dataManagementPlan_long.tex
Updates from Overleaf

The project will generate software modules appropriate for use in behavioural neuroscience experiments in animal models, and other types of study that involve a combination of reactive experimental (particularly behavioural) control and streamed data acquisition.

@@

Software modules will comprise source files in various programming languages: C\#, Python, MATLAB and R.  These will be accompanied by extensive documentation (discussed below).

**NEED**: Approximate lines of code??? (order of magnitude is fine)

**NEED**: testing files, makefiles?, other stuff,

@@

Software will be developed within the git distributed version control system, with public repositories made available within a specially created organisation at github.com.

Copies of all repositories will be maintained within the Sainsbury Wellcome Centre data storage system at UCL.  **NEED**: short/medium term backup protocols from John Pelan's DMP

% management
Package installation and updates are supported by standalone package management
software included within the core Bonsai distribution and hosted in a shared
centralised and curated repository. This ensures that bugfixes and enhancements
are easily available and, importantly, allows for reproducible deployment of
user environments to ensure working experiments are not disrupted by system
changes.
%
Large-scale experimental projects such as the International Brain Laboratory
have leveraged this capability of Bonsai to standardise data acquisition and
control software across dozens of laboratories around the world.
%
Community technology previews are made available through the package manager,
letting early adopters opt-in to the latest features before they are generally
available. This allows the community to report issues with new features on
GitHub, thereby distributing bug reporting and triage, and improving the
quality of stable releases.

The core compiler and IDE is extensively tested, and uses a strong type system
to greatly reduce bugs at compile-time. Best practices for leveraging this
infrastructure for building new packages are in place. The software engineers
recruited to this project will work with our collaborators at NeuroGEARS Ltd,
to strengthen and maintain these practices. Once fully familiar with the Bonsai
core architecture they will participate in review of pull requests and other
code contributions.

@@

% documentation
Current Bonsai documentation, including learning materials and all relevant community
links are publicly accessible at \url{bonsai-rx.org}.

Documentation will be created for the new modules produced in this project that meets the existing standards, and will be integrated into the Bonsai documentation framework.

% The fact that Bonsai is distributed as an open source project
% (\url{https://bonsai-rx.org/}), that is has extensive online documentation
% (\url{https://bonsai-rx.org/docs/}) and a vibrant user discussion forum
% (\url{https://groups.google.com/g/bonsai-users}), and that is integrated with
% many other open-source projects (Figure~1A and~1B, \emph{Case for Support})
% enables the use of the Bonsai ecosystem by the general public.

@@

\section{Data sharing and access}
\subsection{Where will data be shared?}
% Identify any data repository (-ies) that are, or will be, entrusted with storing, curating and/or sharing data from your study, where they exist for particular disciplinary domains or data types.
Software modules and documentation will be available through the github.com service, with links from the Bonsai project site (https://bonsai-rx.org) and available through the Bonsai system's package management interface.

**NEED**: github org and status?

Repositories will be mirrored within the Sainsbury Wellcome Centre scientific computing system.  **NEED**: long-term backup from John's DMP

@@

Release management: repositories will be public throughout, but marked as developmental.  Alpha-level code will be available but not publicised until after extensive testing.  Will be integrated into package management and advertised once it achieves release quality.

@@

Package management
Bonsai website
Community fora
Workshops
Paper?
 A forum where Bonsai
users daily exchange questions and answers can be accessed at
\url{https://groups.google.com/g/bonsai-users}. Currently more than 700 users
contribute to this forum (Figure~1f, \emph{Case for Support}).

@@

This is an open source software project.  Code will be distributed under **NEED** licence, and thus will be freely reusable and modifiable.

@@

>>>>

The project will generate software modules appropriate for use in behavioural neuroscience experiments using animal models, and other types of study that involve a combination of reactive experimental (particularly behavioural) control and streamed data acquisition.

@@

Software modules will comprise source files in various programming languages: C\#, Python, MATLAB and R, along with supporting files for version control, compilation, testing, packaging and documentation.

@@

Software will be developed within the git distributed version control
system, with public repositories made available within a specially
created organisation at github.com and linked from \url{github.com/bonsai-rx}.

Mirrors of all repositories will be maintained within the Sainsbury Wellcome Centre data storage system at UCL.

% management
Package installation and updates are supported by standalone package management
software included within the core Bonsai distribution and hosted in a shared
centralised and curated repository. This ensures that bugfixes and enhancements
are easily available and, importantly, allows for reproducible deployment of
user environments to guarantee existing experiments are not disrupted by system
changes.
%
Large-scale experimental projects such as the International Brain Laboratory
have leveraged this capability of Bonsai to standardise data acquisition and
control software across dozens of laboratories around the world.
%
Community technology previews are made available through the package manager,
letting early adopters opt-in to the latest features before they are generally
available. This allows the community to report issues with new features on
GitHub, thereby distributing bug reporting and triage, and improving the
quality of stable releases.

The core compiler and the Integrated Development Environment (IDE) is extensively tested, and uses a strongly typed language
that greatly reduce bugs at compile-time. Best practices for leveraging this
infrastructure for building new packages are in place. The software engineers
recruited to this project will work with our collaborators at NeuroGEARS Ltd,
to strengthen and maintain these practices. Once fully familiar with the Bonsai
core architecture they will participate in review of pull requests and other
code contributions.

@@

% documentation
Bonsai documentation, including learning materials and all relevant community
links are publicly accessible at \url{bonsai-rx.org}.
% A forum where Bonsai users daily exchange questions and answers can be accessed at the public
% repository \url{https://groups.google.com/g/bonsai-users}. Currently more than
% 700 users contribute to this forum (Figure~1f, \emph{Case for Support}).
Documentation will be created for the new modules produced in this project that
meets the existing standards, and will be integrated into the Bonsai
documentation framework.

Classes, methods and functions will be documented in the code, and automatic
documentation tools (e.g., Sphinx for Python code, or Sphinx for .Net for C\#)
will be used to automatically generate API documentation. This documentation
will be complemented with class and interaction diagrams, documenting the
structure and dynamics of the application objects.

@@

Persistent storage and long term preservation will be achieved using
Zenodo [\url{https://zenodo.org/}], which is well integrated with
GitHub.

\section{Data sharing and access}
\subsection{Where will data be shared?}
% Identify any data repository (-ies) that are, or will be, entrusted with storing, curating and/or sharing data from your study, where they exist for particular disciplinary domains or data types.
Software modules will be available on Github and on Zenodo for long-term
preservation, with links from the Bonsai project site
(\url{https://bonsai-rx.org}) and available through the Bonsai system's package
management interface. Tutorials on how to use the added machine learning
functionality will be available at the Bonsai project site
(\url{https://bonsai-rx.org}) and automatically generated API documentation for
this functionality will be available through ReadTheDocs [\url{https://readthedocs.org/}].

@@

% Release management: repositories will be public throughout, but marked as developmental.  Alpha-level code will be available but not publicised until after extensive testing.  Will be integrated into package management and advertised once it achieves release quality.

Modules will be made available once they have passed rigorous internal quality checks.
Software will be automatically tested during development with a comprehensive set of unit test and integration tests, implemented as  GitHub Actions.  Once complete, each 
module will be tested by a group of beta adopters (largely drawn from within the SWC to enable rapid feedback and debugging).  Project outputs will be accepted when all tests run successfully and no major functional problems are reported in beta tests.

Release management: we will create a new release after each of the machine
learning functionalities (described in Section~2.3 of the Case for Support) is
integrated into the Bonsai ecosystem following the above method to accept a
project output. Credentials for release will be granted initially by Dr.
Gonçalo Lopes, Head Engineer at Neurogears and the current curator of the
Bonsai official package feed, but might be expanded to other team members after
training.  Acceptance criteria for release will be a group decision between
Dr.~Lopes, Dr.~Rapela, Prof.~Sahani and Prof.~Mrsic-Flogel.

@@

Bonsai is already a widely adopted software ecosystem.  Community users will engage with the development of new capabilities in this project through planned workshops and the Bonsai users forum [\url{https://groups.google.com/g/bonsai-users}].

On release, modules will be announced on the Bonsai website and available for installation through the package management system.

We anticipate that incorporation of these new capabilities, associated workshops, announcements, and possible future publications will drive further adoption of Bonsai in behavioural systems neuroscience laboratories worldwide.

@@

Several factors will contribute to the reusability of new functionality
incorporated into Bonsai. Code implementing this functionality will be
available in Github and Zenodo, under the permissive MIT license. Emphasis on standard and
uniform code style, achieved with the help of code review and paired
programming, will generate readable code, easier to read and extend by third
parties. Automated build methods will simplify code building and encourage
modification by new developers. Test cases with large coverage will motivate
software developers to modify Bonsai, without the risk of introducing
undetected bugs into the software. Carefully designed software, with simple and
clear object interfaces, will simplify the reuse of the proposed software.

@@

\section{Capabilities}

The project collaborator, NeuroGEARS Ltd, and its lead, Dr.~Gon\c{c}alo Lopes, have extended experience of software management and distribution, and will support our efforts.  In addition we will benefit from advice from UCL's recently established Centre for Advanced Research Computing.

\section{Responsibilities}

Drs.~Rapela (team member) and Lopes (collaborator) will jointly take responsibility for technical quality of version control, testing, documentation, and distribution.  Oversight will be provided by Profs.~Sahani (co-I) and Mrsic-Flogel (PI) along with a planned scientific advisory board committee that will include international experts.

\section{Relevant institutional, departmental or study policies on data sharing and data security}

Not applicable.

\section{Author}

The plan was prepared jointly by the team.  Enquiries should be directed to the PI and Co-I.
